---
title: Voting Statistics
format:
  dashboard:
    orientation: columns
execute:
  echo: false
  warning: false
  error: false
  message: false
  freeze: false
---

```{r}
library(tidyverse)
library(scales)
library(DT)
source("functions.R")

data = readxl::read_xlsx(here::here("whiteboardpollresults.xlsx")) %>%
  filter(status == "complete") %>%
  mutate(
    across(everything(), ~ replace_na(.x, 0)),
    votes = hq_votes + rtp_votes) %>%
  group_by(series_no, series, date, question) %>%
  mutate(
    N = sum(votes),
    share = round( (votes/N) * 100, 2)) %>%
  ungroup()

pollsasked = length(unique(data$question))
votescounted = sum(data$votes)

mostvoted = unique((data %>% filter(votes == max(data$votes, na.rm = TRUE)))$option)
mostvotes = max(data$votes, na.rm = TRUE)
mostvotes_poll = pull(filter(data, votes == mostvotes), series)
mostvotes_question = pull(filter(data, votes == mostvotes), question)

mostunanimous = unique((data %>% filter(share == max(data$share, na.rm = TRUE)))$option)
mostunanimous_pct = max(data$share, na.rm = TRUE)
mostunanimous_poll = pull(filter(data, share == mostunanimous_pct), series)
mostunanimous_question = pull(filter(data, share == mostunanimous_pct), question)

# Format one or more tied rows for an extreme-value superlative. When only one
# row hits the extreme, the caller builds its own single-row sentence; this is
# only needed to render every tied row (poll + question + option) once each,
# instead of the pull()-per-column approach silently concatenating mismatched
# vectors when more than one row shares the extreme value.
format_ties = function(rows) {
  rows %>%
    mutate(label = sprintf("%s (%s - %s)", option, series, question)) %>%
    pull(label) %>%
    paste(collapse = "; ")
}

fewestvotes = min(data$votes, na.rm = TRUE)
fewest_rows = data %>% filter(votes == fewestvotes) %>% distinct(series, question, option)
leastvoted = unique(fewest_rows$option)
fewest_tied = nrow(fewest_rows) > 1
fewest_detail = format_ties(fewest_rows)

nonzero_votes = data %>% filter(votes > 0)
fewestvotes_nonzero = min(nonzero_votes$votes, na.rm = TRUE)
fewest_nonzero_rows = nonzero_votes %>% filter(votes == fewestvotes_nonzero) %>% distinct(series, question, option)
fewest_nonzero_tied = nrow(fewest_nonzero_rows) > 1
fewest_nonzero_detail = format_ties(fewest_nonzero_rows)

leastunanimous_pct = min(data$share, na.rm = TRUE)
leastunanimous_rows = data %>% filter(share == leastunanimous_pct) %>% distinct(series, question, option)
leastunanimous = unique(leastunanimous_rows$option)
leastunanimous_tied = nrow(leastunanimous_rows) > 1
leastunanimous_detail = format_ties(leastunanimous_rows)

leastunanimous_pct_nonzero = min(nonzero_votes$share, na.rm = TRUE)
leastunanimous_nonzero_rows = nonzero_votes %>% filter(share == leastunanimous_pct_nonzero) %>% distinct(series, question, option)
leastunanimous_nonzero_tied = nrow(leastunanimous_nonzero_rows) > 1
leastunanimous_nonzero_detail = format_ties(leastunanimous_nonzero_rows)

# Raw vote counts (and raw shares) aren't comparable across polls with a different number of options -- a 7-way poll mechanically splits votes thinner than a 2-way poll, so its runt option will look "lower" than a 2-way poll's runt even when neither is unusually spurned. Normalizing each option's share against the chance level for that question (1/k, the same baseline drawn in multi_option_ci_plot()) makes "how badly did this option underperform" comparable no matter how many options the question had.

chance_data = data %>%
  group_by(series_no, series, question) %>%
  mutate(k = n(), chance_share = 100 / k, ratio_to_chance = share / chance_share) %>%
  ungroup()

format_spurned = function(rows) {
  rows %>%
    mutate(label = sprintf("%s got just %sx its fair share (%d of %d votes across %d options; %s - %s)",
                            option, round(ratio_to_chance, 2), votes, N, k, series, question)) %>%
    pull(label) %>%
    paste(collapse = "; ")
}

nonzero_chance = chance_data %>% filter(votes > 0)
most_spurned_ratio = min(nonzero_chance$ratio_to_chance, na.rm = TRUE)
most_spurned_rows = nonzero_chance %>%
  filter(ratio_to_chance == most_spurned_ratio) %>%
  distinct(series, question, option, votes, N, k, ratio_to_chance)
most_spurned_tied = nrow(most_spurned_rows) > 1
most_spurned_text = if (most_spurned_tied) {
  sprintf("tied across %d options: %s", nrow(most_spurned_rows), format_spurned(most_spurned_rows))
} else {
  format_spurned(most_spurned_rows)
}

# --- HQ vs RTP alignment ---------------------------------------------------
# Leader-only HQ/RTP comparison for every two-option question, across every
# poll RTP has voted in -- the same method each blog post uses for its
# "this week" Alignment Index. The cumulative_alignment() figure below uses a
# different (all-options) method, exactly like each blog post shows both
# side by side, so we keep that same pairing here.
hq_rtp_all = data %>%
  group_by(series_no, series, question) %>%
  filter(n() == 2, sum(rtp_votes) > 0) %>%
  group_modify(~ {
    df    = .x %>% arrange(desc(votes))
    top   = df %>% slice(1)
    hq_N  = sum(df$hq_votes)
    rtp_N = sum(df$rtp_votes)
    pt = suppressWarnings(prop.test(x = c(top$hq_votes, top$rtp_votes), n = c(hq_N, rtp_N)))
    tibble(
      option    = top$option,
      hq_share  = top$hq_votes / hq_N,
      rtp_share = top$rtp_votes / rtp_N,
      gap       = (top$hq_votes / hq_N) - (top$rtp_votes / rtp_N),
      p_value   = pt$p.value
    )
  }) %>%
  ungroup()

alignment_by_poll = hq_rtp_all %>%
  group_by(series_no, series) %>%
  summarize(alignment = cor(hq_share, rtp_share), .groups = "drop") %>%
  mutate(method = "two-option (leader only)") %>%
  arrange(series_no)

# A poll where every question has more than two options (e.g. Week Daze) has
# no leader-only pair to correlate, so it never shows up in alignment_by_poll
# above -- but each such poll's post still reports its own Alignment Index,
# using every option's HQ/RTP share instead of just the leader's (the same
# all-options method cumulative_alignment() uses). Compute that per poll here
# so those polls can appear in the trend chart too, clearly marked as a
# different method rather than silently mixed in as if it were the same
# statistic.
multi_option_alignment_by_poll = data %>%
  group_by(series_no, series, question) %>%
  filter(n() > 2, sum(rtp_votes) > 0) %>%
  mutate(hq_share = hq_votes / sum(hq_votes), rtp_share = rtp_votes / sum(rtp_votes)) %>%
  ungroup() %>%
  group_by(series_no, series) %>%
  summarize(alignment = cor(hq_share, rtp_share), .groups = "drop") %>%
  mutate(method = "multi-option (all shares)") %>%
  arrange(series_no)

alignment_by_poll_all = bind_rows(alignment_by_poll, multi_option_alignment_by_poll) %>%
  arrange(series_no)

latest_complete_series = max(data$series_no)
cumulative_alignment_index = cumulative_alignment(latest_complete_series)
total_flips = flip_count_so_far(latest_complete_series)
latest_complete_series_name = data %>% filter(series_no == latest_complete_series) %>% pull(series) %>% unique()

# Detail behind total_flips above: every question (any number of options)
# where HQ's and RTP's own winners differed, with each office's winning
# margin -- how far ahead their pick was of their own runner-up -- rather
# than any comparison across offices, since the two offices didn't pick the
# same option to begin with. total_margin_pts (the two margins summed) is a
# rough "how emphatic was the standoff" score: high when both offices were
# individually decisive about their own, differing pick, not a measure of
# how far apart their support for the same option was (there isn't one here).
different_crowns = data %>%
  filter(series_no >= 12, series_no <= latest_complete_series) %>%
  group_by(series_no, series, question) %>%
  filter(sum(rtp_votes) > 0) %>%
  group_modify(~ {
    hq_sorted  = .x %>% arrange(desc(hq_votes))
    rtp_sorted = .x %>% arrange(desc(rtp_votes))
    tibble(
      hq_winner      = hq_sorted$option[1],
      hq_margin_pts  = (hq_sorted$hq_votes[1] - hq_sorted$hq_votes[2]) / sum(.x$hq_votes) * 100,
      rtp_winner     = rtp_sorted$option[1],
      rtp_margin_pts = (rtp_sorted$rtp_votes[1] - rtp_sorted$rtp_votes[2]) / sum(.x$rtp_votes) * 100
    )
  }) %>%
  ungroup() %>%
  filter(hq_winner != rtp_winner) %>%
  mutate(total_margin_pts = hq_margin_pts + rtp_margin_pts) %>%
  arrange(desc(total_margin_pts))

# How many option-level data points actually feed the cumulative correlation
# -- mirrors cumulative_alignment()'s own filtering -- so the explanation
# below can cite a real, current number instead of a stale hard-coded one.
alignment_pool_n = data %>%
  filter(series_no >= 12, series_no <= latest_complete_series) %>%
  group_by(series_no, question) %>%
  filter(sum(rtp_votes) > 0) %>%
  ungroup() %>%
  nrow()

most_aligned_poll  = alignment_by_poll_all %>% arrange(desc(alignment)) %>% slice(1)
least_aligned_poll = alignment_by_poll_all %>% arrange(alignment) %>% slice(1)
biggest_gap        = hq_rtp_all %>% arrange(desc(abs(gap))) %>% slice(1)

# The gap with the biggest raw margin isn't necessarily the most statistically
# significant one (significance depends on sample size too), so track them
# separately and generate the summary sentence from whatever's actually true
# right now instead of a claim that was only true when it was written.
significant_gaps     = hq_rtp_all %>% filter(p_value < 0.05) %>% arrange(p_value)
n_significant_gaps   = nrow(significant_gaps)
most_significant_gap = significant_gaps %>% slice(1)

significance_note = if (n_significant_gaps == 0) {
  "No HQ/RTP gap has cleared statistical significance yet."
} else if (biggest_gap$p_value >= 0.05) {
  sprintf(
    "It hasn't cleared statistical significance itself -- that distinction currently belongs to %s (%s, p = %s).",
    most_significant_gap$question[1], most_significant_gap$series[1], signif(most_significant_gap$p_value[1], 2)
  )
} else if (n_significant_gaps == 1) {
  "It's also the only office gap that's cleared statistical significance so far."
} else {
  sprintf("It's one of %d office gaps that have cleared statistical significance so far.", n_significant_gaps)
}

# --- Decisiveness across every two-option question ever asked --------------
two_option_all = data %>%
  group_by(series_no, series, question) %>%
  filter(n() == 2) %>%
  group_modify(~ {
    Nq = unique(.x$N)
    ordered = .x %>% arrange(desc(votes))
    lead  = ordered %>% slice(1)
    trail = ordered %>% slice(2)
    bt = binom.test(lead$votes, Nq, p = 0.5)
    tibble(
      N = Nq,
      leader = lead$option, leader_votes = lead$votes,
      trailer = trail$option, trailer_votes = trail$votes,
      p_value = bt$p.value,
      is_tie = lead$votes == trail$votes
    )
  }) %>%
  ungroup()

most_decisive = two_option_all %>% filter(!is_tie) %>% arrange(p_value) %>% slice(1)

# Ties for the *lead*, across every question ever asked. A two-option tie
# (dead-even split) is definitionally the least decisive result possible --
# that's the "closest vote ever" framing below. A multi-option tie is a
# different animal: two or more options can share the top spot while the
# question as a whole is still extremely decisive (everything else was far
# behind), so it doesn't belong in that same "closest" claim -- it just goes
# in the general tie tracker.
two_option_ties = two_option_all %>%
  filter(is_tie) %>%
  transmute(
    series_no, series, question,
    k = 2L, n_tied = 2L,
    tied_votes = leader_votes,
    tied_options = paste(leader, trailer, sep = " vs ")
  ) %>%
  arrange(series_no)

multi_option_all = data %>%
  group_by(series_no, series, question) %>%
  filter(n() > 2) %>%
  group_modify(~ {
    max_votes = max(.x$votes)
    top_rows  = .x %>% filter(votes == max_votes) %>% arrange(option)
    tibble(
      k            = nrow(.x),
      n_tied       = nrow(top_rows),
      tied_votes   = max_votes,
      tied_options = paste(top_rows$option, collapse = " vs ")
    )
  }) %>%
  ungroup()

multi_option_ties = multi_option_all %>%
  filter(n_tied > 1) %>%
  select(series_no, series, question, k, n_tied, tied_votes, tied_options)

all_ties = bind_rows(two_option_ties, multi_option_ties) %>% arrange(series_no)

# --- RTP participation -----------------------------------------------------
rtp_era = data %>% filter(series_no >= 12)
rtp_total_votes          = sum(rtp_era$rtp_votes)
hq_total_votes_since_rtp = sum(rtp_era$hq_votes)
rtp_overall_share        = rtp_total_votes / (rtp_total_votes + hq_total_votes_since_rtp)

participation_by_poll = rtp_era %>%
  group_by(series_no, series) %>%
  summarize(hq_total = sum(hq_votes), rtp_total = sum(rtp_votes), .groups = "drop") %>%
  mutate(rtp_share_of_poll = rtp_total / (hq_total + rtp_total)) %>%
  arrange(series_no)
```

# Overview

## Overview

### Totals

***`r pollsasked`*** polls conducted & ***`r votescounted`*** votes counted

### Superlatives

[Highest]{style="color: #4B824F"} vote count: ***`r mostvotes`*** for ***`r mostvoted`***
(Poll: `r mostvotes_poll`, Question: `r mostvotes_question`)

[Most decisive vote ever]{style="color: #4B824F"}: ***`r most_decisive$leader`*** over ***`r most_decisive$trailer`***, `r most_decisive$leader_votes`–`r most_decisive$trailer_votes`
(Poll: `r most_decisive$series`, Question: `r most_decisive$question`; p = `r signif(most_decisive$p_value, 2)`)

[Most unanimous]{style="color: #4B824F"} vote: ***`r mostunanimous`*** with a *resounding* ***`r mostunanimous_pct`*** % of the votes
(Poll: `r mostunanimous_poll`, Question: `r mostunanimous_question`)

<!-- [Lowest]{style="color: #824B4B"} vote count: ***`r fewestvotes`*** `r if (fewest_tied) sprintf("- tied across %d options: %s", nrow(fewest_rows), fewest_detail) else sprintf("for ***%s*** (Poll: %s, Question: %s)", leastvoted, fewest_rows$series[1], fewest_rows$question[1])` -->

<!-- [Lowest (non-zero)]{style="color: #824B4B"} vote count: ***`r fewestvotes_nonzero`*** `r if (fewest_nonzero_tied) sprintf("- tied across %d options: %s", nrow(fewest_nonzero_rows), fewest_nonzero_detail) else sprintf("for ***%s*** (Poll: %s, Question: %s)", fewest_nonzero_rows$option[1], fewest_nonzero_rows$series[1], fewest_nonzero_rows$question[1])` -->

<!-- [Least unanimous]{style="color: #824B4B"} vote: `r if (leastunanimous_tied) sprintf("***%s%%*** of the votes - tied across %d options: %s", leastunanimous_pct, nrow(leastunanimous_rows), leastunanimous_detail) else sprintf("***%s*** with ***%s%%*** of the votes (Poll: %s, Question: %s)", leastunanimous, leastunanimous_pct, leastunanimous_rows$series[1], leastunanimous_rows$question[1])` -->

[Least unanimous (non-zero)]{style="color: #824B4B"} vote: `r if (leastunanimous_nonzero_tied) sprintf("***%s%%*** of the votes - tied across %d options: %s", leastunanimous_pct_nonzero, nrow(leastunanimous_nonzero_rows), leastunanimous_nonzero_detail) else sprintf("***%s*** with ***%s%%*** of the votes (Poll: %s, Question: %s)", leastunanimous_nonzero_rows$option[1], leastunanimous_pct_nonzero, leastunanimous_nonzero_rows$series[1], leastunanimous_nonzero_rows$question[1])`

[Most-spurned option ever]{style="color: #824B4B"}: `r most_spurned_text`
<br><span style="font-size: 0.85em; color: gray;">This compares each option's share of votes to the chance level for its own question (1/*k*, where *k* is however many options that question had) — the lowest such ratio, among options that got at least one vote, is comparable no matter how many-way the poll was.</span>

[Closest vote(s) ever]{style="color: #6425B8"}: a ***`r nrow(two_option_ties)`-way tie*** for closest *(closest possible result in a 2-way poll — a dead-even 50/50 split)* — see the tie tracker tab.
<br><span style="font-size: 0.85em; color: gray;">This is scoped to two-option questions only, since a dead-even split is the closest a vote can possibly land. Multi-option ties for the lead are a different kind of tie — tracked separately in the tie tracker — and don't necessarily mean the overall result was close.</span>

# Tie Tracker

::: {.card title="Tie tracker"}

***`r nrow(all_ties)`*** questions have ended in a tie for the lead so far — `r nrow(two_option_ties)` as a dead-even two-option split, and `r nrow(multi_option_ties)` as a shared top spot among more than two options:

```{r}
all_ties %>%
  transmute(
    Poll = series,
    Question = question,
    Split = if_else(
      k == 2,
      paste0(tied_votes, "-", tied_votes, " (", tied_options, ")"),
      paste0(tied_votes, " votes each, tied for 1st of ", k, " options (", tied_options, ")")
    )
  ) %>%
  knitr::kable()
```

:::

# HQ + RTP

## Alignment

### Alignment Index

Cumulative correlation between HQ's and RTP's vote shares, across every option RTP has voted on since it joined the poll at Earth Day:

***`r round(cumulative_alignment_index, 2)`*** cumulative Alignment Index (Earth Day → `r latest_complete_series_name`)

This is a different calculation from the per-poll bars (see Alignment by poll tab to the right), not their average. It's also asking a stricter question than "did the two offices pick the same winner" -- it's asking whether they were equally *enthusiastic* about every option. Each per-poll bar answers that on so few votes that it can swing wildly: Fair Fare reads as negative even though HQ and RTP agreed on every winner, just with slightly different enthusiasm. This number pools every vote from every poll instead, so it's a far more reliable read on how closely the two offices actually vote alike.

### Alignment superlatives

[Most aligned poll]{style="color: #4B824F"}: ***`r most_aligned_poll$series`*** (`r round(most_aligned_poll$alignment, 2)`)

[Least aligned poll]{style="color: #824B4B"}: ***`r least_aligned_poll$series`*** (`r round(least_aligned_poll$alignment, 2)`)

### Flips & gaps

***`r total_flips`*** questions where HQ and RTP would have crowned different winners on their own, since RTP joined the poll.

[Biggest HQ/RTP gap on record]{style="color: #82704B"}: ***`r biggest_gap$question`*** (`r biggest_gap$series`) — HQ `r round(biggest_gap$hq_share*100,1)`% vs RTP `r round(biggest_gap$rtp_share*100,1)`% for *`r biggest_gap$option`* (p = `r signif(biggest_gap$p_value,2)`). `r significance_note`

## Graphs

### Graphs

::: {.panel-tabset}

#### Alignment by poll

The cumulative Alignment Index (see the Alignment section left) pools every option from every poll into one correlation; this instead shows each poll's own alignment on its own, so you can see how much any single poll's read can wander from that cumulative figure (dashed line) -- see the Alignment section above for why that's expected rather than a discrepancy.

```{r}
#| label: fig-alignment-trend
#| fig-cap: "HQ/RTP correlation, one poll at a time. Two-option polls use the leader-only method; polls where every question has more than two options (no single leader pair to compare) use the all-shares method instead -- the same distinction each poll's own post draws."
ggplot(alignment_by_poll_all, aes(x = fct_reorder(series, series_no), y = alignment, fill = method)) +
  geom_col(width = 0.6) +
  geom_hline(yintercept = cumulative_alignment_index, linetype = "dashed", color = "gray40") +
  geom_text(aes(label = round(alignment, 2)), vjust = -0.5, size = 3.5) +
  scale_y_continuous(limits = c(min(0, min(alignment_by_poll_all$alignment) - 0.05), 1)) +
  scale_fill_manual(values = c("two-option (leader only)" = "#4B6B82", "multi-option (all shares)" = "#4B826F"), name = NULL) +
  labs(x = NULL, y = "Alignment (that poll only)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "bottom")
```

#### Participation share

RTP has cast ***`r rtp_total_votes`*** of the ***`r rtp_total_votes + hq_total_votes_since_rtp`*** votes counted since it joined — ***`r round(rtp_overall_share*100,1)`%*** of the total. That share hasn't trended up or down so much as bounced around a fairly consistent range, poll to poll:

```{r}
#| label: fig-rtp-participation
#| fig-cap: "RTP's share of the vote, one poll at a time."
ggplot(participation_by_poll, aes(x = fct_reorder(series, series_no), y = rtp_share_of_poll)) +
  geom_col(fill = "#82704B", width = 0.6) +
  geom_hline(yintercept = rtp_overall_share, linetype = "dashed", color = "gray40") +
  geom_text(aes(label = percent(rtp_share_of_poll, accuracy = 1)), vjust = -0.5, size = 3.5) +
  scale_y_continuous(limits = c(0, 0.6), labels = percent_format()) +
  labs(x = NULL, y = "RTP share of that poll's votes") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

#### Different crowns

Every question where HQ and RTP would have crowned different winners on their own, since RTP joined the poll -- each office's pick and how far ahead of their own runner-up it was, sorted by the two margins combined (a rough read on how emphatic the standoff was, not on how far apart the offices' support for the same option was -- they didn't pick the same option here).

```{r}
#| label: tbl-different-crowns
different_crowns %>%
  transmute(
    Poll = series,
    Question = question,
    `HQ pick` = sprintf("%s (+%.1f pts)", hq_winner, hq_margin_pts),
    `RTP pick` = sprintf("%s (+%.1f pts)", rtp_winner, rtp_margin_pts),
    `Combined margin` = sprintf("%.1f pts", total_margin_pts)
  ) %>%
  knitr::kable()
```

:::

