---
title: Voting Statistics
format:
  dashboard:
    orientation: columns
execute:
  echo: false
  warning: false
  error: false
  message: false
---

```{r}
library(tidyverse)
library(scales)
library(DT)
source("functions.R")

data = readxl::read_xlsx(here::here("whiteboardpollresults.xlsx")) %>%
  filter(status == "complete") %>%
  mutate(
    across(everything(), ~ replace_na(.x, 0)),
    votes = hq_votes + rtp_votes) %>%
  group_by(series_no, series, date, question) %>%
  mutate(
    N = sum(votes),
    share = round( (votes/N) * 100, 2)) %>%
  ungroup()

pollsasked = length(unique(data$question))
votescounted = sum(data$votes)

mostvoted = unique((data %>% filter(votes == max(data$votes, na.rm = TRUE)))$option)
mostvotes = max(data$votes, na.rm = TRUE)
mostvotes_poll = pull(filter(data, votes == mostvotes), series)
mostvotes_question = pull(filter(data, votes == mostvotes), question)

mostunanimous = unique((data %>% filter(share == max(data$share, na.rm = TRUE)))$option)
mostunanimous_pct = max(data$share, na.rm = TRUE)
mostunanimous_poll = pull(filter(data, share == mostunanimous_pct), series)
mostunanimous_question = pull(filter(data, share == mostunanimous_pct), question)

leastvoted = unique((data %>% filter(votes == min(data$votes, na.rm = TRUE)))$option)
fewestvotes = min(data$votes, na.rm = TRUE)
fewest_poll = pull(filter(data, votes == fewestvotes), series)
fewest_question = pull(filter(data, votes == fewestvotes), question)

leastunanimous = unique((data %>% filter(share == min(data$share, na.rm = TRUE)))$option)
leastunanimous_pct = min(data$share, na.rm = TRUE)
leastunanimous_poll = pull(filter(data, share == leastunanimous_pct), series)
leastunanimous_question = pull(filter(data, share == leastunanimous_pct), question)

# --- HQ vs RTP alignment ---------------------------------------------------
# Leader-only HQ/RTP comparison for every two-option question, across every
# poll RTP has voted in -- the same method each blog post uses for its
# "this week" Alignment Index. The cumulative_alignment() figure below uses a
# different (all-options) method, exactly like each blog post shows both
# side by side, so we keep that same pairing here.
hq_rtp_all = data %>%
  group_by(series_no, series, question) %>%
  filter(n() == 2, sum(rtp_votes) > 0) %>%
  group_modify(~ {
    df    = .x %>% arrange(desc(votes))
    top   = df %>% slice(1)
    hq_N  = sum(df$hq_votes)
    rtp_N = sum(df$rtp_votes)
    pt = suppressWarnings(prop.test(x = c(top$hq_votes, top$rtp_votes), n = c(hq_N, rtp_N)))
    tibble(
      option    = top$option,
      hq_share  = top$hq_votes / hq_N,
      rtp_share = top$rtp_votes / rtp_N,
      gap       = (top$hq_votes / hq_N) - (top$rtp_votes / rtp_N),
      p_value   = pt$p.value
    )
  }) %>%
  ungroup()

alignment_by_poll = hq_rtp_all %>%
  group_by(series_no, series) %>%
  summarize(alignment = cor(hq_share, rtp_share), .groups = "drop") %>%
  arrange(series_no)

latest_complete_series = max(data$series_no)
cumulative_alignment_index = cumulative_alignment(latest_complete_series)
total_flips = flip_count_so_far(latest_complete_series)

most_aligned_poll  = alignment_by_poll %>% arrange(desc(alignment)) %>% slice(1)
least_aligned_poll = alignment_by_poll %>% arrange(alignment) %>% slice(1)
biggest_gap        = hq_rtp_all %>% arrange(desc(abs(gap))) %>% slice(1)

# --- Decisiveness across every two-option question ever asked --------------
two_option_all = data %>%
  group_by(series_no, series, question) %>%
  filter(n() == 2) %>%
  group_modify(~ {
    Nq = unique(.x$N)
    ordered = .x %>% arrange(desc(votes))
    lead  = ordered %>% slice(1)
    trail = ordered %>% slice(2)
    bt = binom.test(lead$votes, Nq, p = 0.5)
    tibble(
      N = Nq,
      leader = lead$option, leader_votes = lead$votes,
      trailer = trail$option, trailer_votes = trail$votes,
      p_value = bt$p.value,
      is_tie = lead$votes == trail$votes
    )
  }) %>%
  ungroup()

most_decisive = two_option_all %>% filter(!is_tie) %>% arrange(p_value) %>% slice(1)
ties = two_option_all %>% filter(is_tie) %>% arrange(series_no)

# --- RTP participation -----------------------------------------------------
rtp_era = data %>% filter(series_no >= 12)
rtp_total_votes          = sum(rtp_era$rtp_votes)
hq_total_votes_since_rtp = sum(rtp_era$hq_votes)
rtp_overall_share        = rtp_total_votes / (rtp_total_votes + hq_total_votes_since_rtp)

participation_by_poll = rtp_era %>%
  group_by(series_no, series) %>%
  summarize(hq_total = sum(hq_votes), rtp_total = sum(rtp_votes), .groups = "drop") %>%
  mutate(rtp_share_of_poll = rtp_total / (hq_total + rtp_total)) %>%
  arrange(series_no)
```

# Overview

## Overview

### Totals

***`r pollsasked`*** polls conducted & ***`r votescounted`*** votes counted

### Superlatives

[Highest]{style="color: #4B824F"} vote count: ***`r mostvotes`*** for ***`r mostvoted`***
(Poll: `r mostvotes_poll`, Question: `r mostvotes_question`)

[Most decisive vote ever]{style="color: #4B824F"}: ***`r most_decisive$leader`*** over ***`r most_decisive$trailer`***, `r most_decisive$leader_votes`–`r most_decisive$trailer_votes`
(Poll: `r most_decisive$series`, Question: `r most_decisive$question`; p = `r signif(most_decisive$p_value, 2)`)

[Most unanimous]{style="color: #4B824F"} vote: ***`r mostunanimous`*** with a *resounding* ***`r mostunanimous_pct`*** % of the votes
(Poll: `r mostunanimous_poll`, Question: `r mostunanimous_question`)

[Lowest]{style="color: #824B4B"} vote count: ***`r fewestvotes`*** for ***`r leastvoted`***
(Poll: `r fewest_poll`, Question: `r fewest_question`)

[Least unanimous]{style="color: #824B4B"} vote: ***`r leastunanimous`*** with ***`r leastunanimous_pct`*** % of the votes
(Poll: `r leastunanimous_poll`, Question: `r leastunanimous_question`)

[Closest vote(s) ever]{style="color: #6425B8"}: a ***`r nrow(ties)`-way tie*** for closest — see the tie tracker for the full list.

::: {.card title="Tie tracker"}

***`r nrow(ties)`*** questions have ended in a dead-even split so far:

```{r}
ties %>%
  transmute(
    Poll = series,
    Question = question,
    Split = paste0(leader_votes, "-", trailer_votes, " (", leader, " vs ", trailer, ")")
  ) %>%
  knitr::kable()
```

:::

# HQ + RTP

## Alignment

### Alignment Index

Cumulative correlation between HQ's and RTP's vote shares, across every option RTP has voted on since it joined the poll at Earth Day.

***`r round(cumulative_alignment_index, 2)`*** cumulative Alignment Index (Earth Day → `r tail(alignment_by_poll$series, 1)`)

### Alignment by poll

```{r}
#| label: fig-alignment-trend
#| fig-cap: "Leader-only HQ/RTP correlation, one poll at a time."
ggplot(alignment_by_poll, aes(x = fct_reorder(series, series_no), y = alignment)) +
  geom_col(fill = "#4B6B82", width = 0.6) +
  geom_hline(yintercept = cumulative_alignment_index, linetype = "dashed", color = "gray40") +
  geom_text(aes(label = round(alignment, 2)), vjust = -0.5, size = 3.5) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = NULL, y = "Alignment (that poll only)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

### Alignment superlatives

[Most aligned poll]{style="color: #4B824F"}: ***`r most_aligned_poll$series`*** (`r round(most_aligned_poll$alignment, 2)`)

[Least aligned poll]{style="color: #824B4B"}: ***`r least_aligned_poll$series`*** (`r round(least_aligned_poll$alignment, 2)`)

### Flips & gaps

***`r total_flips`*** questions where HQ and RTP would have crowned different winners on their own, since RTP joined the poll.

[Biggest HQ/RTP gap on record]{style="color: #82704B"}: ***`r biggest_gap$question`*** (`r biggest_gap$series`) — HQ `r round(biggest_gap$hq_share*100,1)`% vs RTP `r round(biggest_gap$rtp_share*100,1)`% for *`r biggest_gap$option`* (p = `r signif(biggest_gap$p_value,2)`). It's also the only office gap that's cleared statistical significance so far.

## Participation

::: {.card title="RTP participation"}

RTP has cast ***`r rtp_total_votes`*** of the ***`r rtp_total_votes + hq_total_votes_since_rtp`*** votes counted since it joined — ***`r round(rtp_overall_share*100,1)`%*** of the total. That share hasn't trended up or down so much as bounced around a fairly consistent range, poll to poll:

```{r}
#| label: fig-rtp-participation
#| fig-cap: "RTP's share of the vote, one poll at a time."
ggplot(participation_by_poll, aes(x = fct_reorder(series, series_no), y = rtp_share_of_poll)) +
  geom_col(fill = "#82704B", width = 0.6) +
  geom_hline(yintercept = rtp_overall_share, linetype = "dashed", color = "gray40") +
  geom_text(aes(label = percent(rtp_share_of_poll, accuracy = 1)), vjust = -0.5, size = 3.5) +
  scale_y_continuous(limits = c(0, 0.6), labels = percent_format()) +
  labs(x = NULL, y = "RTP share of that poll's votes") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

:::

